---
title: "The Brothers Karamazov Data Package"
author: "Casey Crary"
date: "2024-02-19"
date-format: iso
format: gfm
editor: source
---

```{r}
#| label: setup
#| include: false
library(BrothersKaramazov)
library(tidytext)
library(tidyr)
library(dplyr)
library(ggplot2)
```

This file describes the `BrothersKaramazov` package.

This package allows easy analysis of "The Brothers Karamazov", an extremely influential work in Russian literature.  

The BrothersKaramazov package can be installed by running:

```
devtools::install_github("Casey308/BrothersKaramazov")
library(BrothersKaramazov)
```

```{r}
library(tidyverse)
glimpse(karamazov_lines)
```

Here are some sample analyses:

```{r}
karamazov_words <- karamazov_lines |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by("word")) |>
  count(word, sort = TRUE) |>
  mutate(word = reorder(word, n))

ggplot(data = head(karamazov_words, 10), aes(n, word)) +
  geom_col() +
  labs(x = "Count", y = "Word", title = "Most Common Words in 'The Brothers Karamazov'")
```

We can see from this that the three brothers, Alysoha, Mitya, and Ivan, are all in the most common words in the novel.
It also appears that time is a central theme in the book, as "time," "suddenly," and "day," all relate to time in some way and are among the most prevalent words in the novel.

```{r}
#| warning: false
karamazov_sentiments <- karamazov_lines |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by("word")) |>
  inner_join(get_sentiments("bing"), join_by("word")) |>
  count(chapter, sentiment) |>
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |>
  mutate(sentiment = positive - negative)

ggplot(karamazov_sentiments, aes(chapter, sentiment)) +
  geom_col(show.legend = FALSE) +
  labs(x = "Chapter Number", y = "Bing Sentiment",
       title = "Bing Sentiment of Each Chapter")
```

From this plot we can see that there are only 6 chapters that are more positive than negative based on the Bing lexicon.
This is slightly surprising, as this is considered one of Dostoyevsky's happier works.
